Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2022-01-08 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

Light, specifically the proportion of incident light and photosynthetically active radiation (PAR), decreases from the canopy top to the forest floor, with the profile shape modified by leaf area density, leaf clumping, canopy height, and canopy structure across species and forest types [Fig. `r fig_NEON_vertical`a-d, Supporting Information Figure `r SIfig_NEON_all`, @koike_leaf_2001]. Canopy foliage also influence the quality of light or the spectral characteristics along the gradient that shift from wavelengths in blue (400-500nm) in the top canopy layers to wavelengths in near infra red (700-1000nm) in the lower canopy layers and the forest floor. With red to far red ratio decreasing with height.
This gradient is more pronounced in dense canopies, including broad-leaf and mixed forests (e.g., SCBI, SERC and HARV, Fig. `r fig_NEON_vertical`d), than in open canopy forests, including many conifer forests [e.g., OSBS, PUUM, and WREF, Fig. `r fig_NEON_vertical`d, Supporting Information Figure `r SIfig_NEON_all`, @chazdonPhotosyntheticLightEnvironments1984; @aussenacInteractionsForestStands2000a; @smithSeasonalDroughtrelatedChanges2019; @tymenQuantifyingMicroenvironmentalVariation2017;  @parkerConsequencesEnvironmentalHeterogeneity2019; @bartemucciOverstoryInfluencesLight2006; @baldocchiSeasonalVariationEnergy1997]. 
Variability in the light environment decreases with height, with the shaded understory being highly influenced by sunflecks, or brief increases in solar radiation caused by small canopy gaps and wind-induced canopy movements [@waySunflecksTreesForests2012]. 
Likewise, the mid-canopy experiences a highly variable light environment due to the sun's passage across a dynamically structured canopy surface [@waySunflecksTreesForests2012].
The implication of this light gradient is that upper canopy leaves have higher potential photosynthetic rate (see section 4.2), but also greater propensity towards elevated $T_{leaf}$ (see sections 2.2, 4.2).

>>>>

Light, specifically the proportion of incident light and photosynthetically active radiation (PAR), decreases from the canopy top to the forest floor, with the profile shape modified by leaf area density, leaf clumping, canopy height, and canopy structure across species and forest types [Fig. `r fig_NEON_vertical`a-d, Supporting Information Figure `r SIfig_NEON_all`, @koike_leaf_2001]. Canopy foliage filter light along the gradient, therefore influencing the quality of light or the spectral characteristics. In the top canopy layers wavelengths range from 400-500nm in the blue range, whereas in the lower canopy layers and the forest floor, wavelengths are enriched in near infra red ranges (700-1000nm). With red to far red ratio decreasing with height [de Castro 2000].
This gradient is more pronounced in dense canopies, including broad-leaf and mixed forests (e.g., SCBI, SERC and HARV, Fig. `r fig_NEON_vertical`d), than in open canopy forests, including many conifer forests [e.g., OSBS, PUUM, and WREF, Fig. `r fig_NEON_vertical`d, Supporting Information Figure `r SIfig_NEON_all`, @chazdonPhotosyntheticLightEnvironments1984; @aussenacInteractionsForestStands2000a; @smithSeasonalDroughtrelatedChanges2019; @tymenQuantifyingMicroenvironmentalVariation2017;  @parkerConsequencesEnvironmentalHeterogeneity2019; @bartemucciOverstoryInfluencesLight2006; @baldocchiSeasonalVariationEnergy1997]. 
Variability in the light environment decreases with height, with the shaded understory being highly influenced by sunflecks, or brief increases in solar radiation caused by small canopy gaps and wind-induced canopy movements [@waySunflecksTreesForests2012]. 
Likewise, the mid-canopy experiences a highly variable light environment due to the sun's passage across a dynamically structured canopy surface [@waySunflecksTreesForests2012].
The implication of this light gradient is that upper canopy leaves have higher potential photosynthetic rate (see section 4.2), but also greater propensity towards elevated $T_{leaf}$ (see sections 2.2, 4.2).
